validateHOT - Validate your Holdout Task

### 2023-12-21 README.Rmd
proof reading of manuals, markdown, and vignette - 231221

The goal of `validateHOT` is to validate the results of your validation
task (also known as holdout task) as well as a tool for simulating
markets with the results of your MaxDiff, CBC, and ACBC. This package is
especially relevant for the [Sawtooth
Software](https://sawtoothsoftware.com/) community who would like to
report their analysis in *R* for open science purposes. However, it also
works perfectly with the great package `ChoiceModelR`(Sermas, 2022). Of
course, this package is also useful for practitioners, who would like to
run the analyses in an open source software. The ultimate goal of these
preference measurement techniques is to predict future behavior (Green &
Srinivasan, 1990). Therefore, it is essential for both academics and
practitioners to ensure that the collected data is valid and can also
predict outside tasks that were not included in the estimation of the
utility scores.

Although commercial studies often do not include a validation/holdout
task (Yang et al., 2018), it is highly recommended to do so (Orme, 2015;
Rao, 2014). This validation/holdout task does not only help to check
whether everything went right during data collection but also to
determine the final model. `validatHOT` provides some of the relevant
metrics to test the performance of the data in predicting a
validation/holdout task.

@@

👉🏾 <u>**What you need to provide**</u>: <br> After collecting your data,
and running your initial Hierarchical Bayes models, you can turn to
`validateHOT` and test how good your model predicts choices in the
validation/holdout task. Therefore, you only have to read in your raw
utility scores as well as the actual choice of your validation/ holdout
task. You can use the `merge()` function provided by the *R* base
package to do so (R Core Team, 2023). Afterward, you can read in your
data file and enjoy `validateHOT`. We provide a short tutorial in the
markdown, for a more detailed tutorial, please see the vignette that
comes along with `validateHOT`.

@@

<li>validation metrics mainly reported in preference measurement
studies</li>

@@

<li><code>hitrate()</code>: creates the Hit Rate (correctly predicted
choices) of your validation/holdout task. The output will contain the
percentage of how many choices were correctly predicted, the chance
level in your validation task ($\frac{1}{alternatives}$) in percentage,
the number of correctly predicted participants' choices as well as the
number of observations.</li>

<li><code>kl()</code>: Kullback-Leibler-Divergence measures the
divergence between the actual choice distribution and the predicted
choice distribution (Ding et al., 2011; Drost, 2018). Output provides
both divergence between predicted from observed and observed from
predicted due to the asymmetry of the Kullback-Leibler divergence. If
you specify an optional <code>group</code> argument the output is split
by groups. Currently, you can choose between \eqn{log} and \eqn{log{_2}}
as logarithm base. The default is set to \eqn{log}.</li>

@@

<li><code>medae()</code>: calculates the median absolute error, which is
less affected by outliers compared to the mean absolute error.</li>

@@

<li><code>rmse()</code>: provides the root-mean-squared error of
deviation between predicted and stated choice share.</li>

@@

We also include metrics from machine learning, i.e., the confusion
matrix (e.g., Burger, 2018). For all of the 5 provided functions, you
currently have to include a **none** alternative in your
validation/holdout task. We currently predict, e.g., whether a buy or
no-buy was correctly predicted. Information could be used for
overestimating and underestimating, respectively, of product purchases.
In the following <code>TP</code> stands for true positives,
<code>FP</code> for false positives, <code>TN</code> for true negatives,
and <code>FN</code> for false negatives (Burger, 2018). To translate
this to the `validateHOT` logic, imagine you have a validation/holdout
task with 5 alternatives plus the alternative not to buy any of those
chosen. `validateHOT` now measures whether or not a buy (participant
opts for one of the 5 alternatives) or a no-buy (participant opts for
the no-buy alternative), respectively, is correctly predicted.

Please be aware that for coding, `validateHOT` applies the following
coding of the *buy* and *no-buy* alternatives. Rows refer to the
observed decisions while columns refer to the predicted ones.

|        |     |        |
|--------|:---:|:------:|
|        | Buy | No-buy |
| Buy    | TP  |   FN   |
| No-buy | FP  |   TN   |

@@

<li><code>turf()</code>: <b>T</b>(otal) <b>U</b>(nduplicated)
<b>R</b>(each) and <b>F</b>(requency) is a "product line extension
model" (Miaoulis et al., 1990, p. 29) that helps to find the perfect
product bundle based on the reach (e.g., how many participants consider
buying at least one product of that assortment) and the frequency (how
many products are on average a purchase option). <code>turf()</code>
currently provides both the *threshold* approach (<code>approach =
'thres'</code>; all products that exceed a threshold are considered,
e.g., as purchase option; Chrzan & Orme, 2019, p. 112) and the *first
choice* approach (<code>approach = 'thres'</code>; only product with
highest utility is considered as purchase option; Chrzan & Orme, 2019,
p. 111).</li>

<li><code>freqassort()</code>: Similar to <code>turf()</code>,
<code>freqassort()</code> will give you the averaged frequency, how many
products the participants will choose from your in the function
determined potential assortment. Again, you have to define a
<code>none</code> alternative, because <code>freqassort()</code> uses
the <i>threshold</i> approach, meaning if the utility of one product is
above the utility of <code>none</code>, it is marked as potential
purchase option. While <code>turf()</code> calculates the reach and
frequency for <b>all</b> combinations, you specify the combination you
are interested in <code>freqassort()</code>.</li>

@@

<li><code>marksim()</code>: runs market simulations (either share of
preference or first choice rule), including the standard error, as well
as the lower and upper confidence interval, which is calculated
according to the $mean +/- 1.96 x \frac{sd}{\sqrt(n)}$ (Orme, 2020, p.
94).</li>

@@

### Reestimating raw utilities

<code>validateHOT</code> also provides 3 functions to better interpret
the scores of both (A)CBC and MaxDiff. Currently, 3 functions are
provided, namely:

@@

<li><code>att_imp()</code>: converts the raw utilities of either an ACBC
or CBC into importance scores for each attribute (see, Orme, 2020, pp.
79-81)</li>

<li>

<code>prob_scores()</code>converts the raw utilities of a MaxDiff to
choice probabilities by applying the following procedures:

@@

<li>For anchored MaxDiff:
$\frac{exp^{U_i}}{(exp^{U_i} + a - 1)} * 100 * \frac{1}{a}$ (Chrzan &
Orme, 2019, p. 59).</li>

@@

<li><code>zc_diffs()</code>: rescales the raw logit utilities to make
them comparable across participants (Sawtooth Software Inc, 2023, p.
343).</li>

@@

`validaeHOT` provides 5 data sets that come with the package. These data
sets should help to better understand the way the data needs to be
defined, especially for the `createHOT()` function.

@@

<li><code>CBC_lin</code>: Example data set with raw utilities of a CBC
study conducted in Sawtooth. One attribute was linear coded while the
other attributes are part-worth coded.</li>

@@

## Why `validateHOT`

We are teaching a preference measurement seminar for students. Often
these students did not have any prior experience (or only sparsely) with
*R*. One of the chapters in this class is about model validating the
results obtained and we teach this, of course, in *R*. We want to put a
strong emphasis on open science and providing tools to run the analyses
and provide the code afterwards is essential.

Of course, there are other great packages which are faster in running
(i.e., `Metrics` by Hamner & Frasco, 2018), however, these packages need
some more data wrangling in order to use the appropriate functions with
the raw utilities, which might be a burden or barrier for one or the
other.

Moreover, as Yang et al. (2018) report, commercial studies often do not
use any validation task. Again, the missing experience in *R* could be
one explanation. Since these functions are not always implemented in
other software, this might be one reason why they do not include one
simply because they do not know how to use it correctly. Having a
package to evaluate the validation/holdout task can also be beneficial
from this perspective.

@@

First, we load the package.

@@

Now imagine you included a validation/holdout task with 3 alternatives
plus a no-buy alternative. We specify the `data` argument and as well as
the `id`. Since we also have a *no-buy* alternative in our validation
task, we next specify the `none` argument, otherwise we would have left
it empty. Afterwards, we define each alternative with the argument
`prod.levels`. If we look back at the data frame, we can see that the
first alternative in the holdout task (`c(4, 9, 19)`) is composed of the
following attribute levels `Att1_Lev1`, `Att2_Lev1`, and `Att3_Lev5`.

@@

> ❗ `createHOT()` takes both column names and column indexes. However,
> please be aware, if you include linear coding or piecewise coding, you
> **have** to provide the column indexes for the input in prod.levels.,
> please be aware of this. However, you can easily find out the index by
> either using the `names()` function to see all names of the data
> frame, or if you already know the name of the variable, you can use
> `which()` and `colnames()`, both functions are provided by the base
> package (R Core Team, 2023). For example, if you want to find out the
> column index of `ID`, you could run `which(colnames(CBC) == "ID")` to
> determine the column index of `ID`.

@@

Let us take a glimpse at the output, which shows the participants' total
raw utilities for each of the 3 alternatives that were included in the
validation/holdout task.

@@

In the next step, we would like to see how well our model (from which we
took the raw utilities) predicts the actual choices in the
validation/holdout task. First, we will run the `hitrate()` function. We
specify the `data`, the column names of the alternatives (`opts`;
remember there are 3 alternatives + the *no-buy* alternative), and
finally the actual choice (`choice`).

@@

Finally, let us test, how many participants would at least buy one of
the 3 products, assuming that this is one potential assortment we would
like to offer to our consumers. We will use the `reach()` function. To
specify the bundles we are offering we use the `opts` argument in our
function.

@@

### Example II - CBC with linear coding

In a second example, we again use a *CBC*, however, this time we show
how to use `validateHOT` if one of the variables are linear coded. All
other examples are provided in the accompanied vignette.

@@

Next, we create the validation/holdout task to evaluate it in the next
step. We use a validation/holdout task with 3 alternatives plus the
*no-buy* alternative, just as we did in the previous example. The only
difference to the previous example is that for our model estimation, the
third attribute (`Att3_Lin`) was coded as linear.

Again, we first define data, namely the `id` as well as the `none`
alternative. Next, we define the `prod.levels` for each alternative.
Since we have one linear coded attribute, we have to specify the column
indexes instead of the column names. For example, for the second
alternative which is composed of `Att1_Lev5`, `Att2_Lev4`, and `60`
which is the value that should be interpolated. We tell `createHOT()`
that this value needs to be interpolated by specifying the `coding`
argument. This tells us that the first two attributes (`Att1_Lev5`,
`Att2_Lev4` in the case of alternative 2) are part-worth coded (`0`)
while the third attribute is linear coded (`1`).

To interpolate the value, we have to provide `createHOT()` the
`interpolate.levels`. These **need** to be the same as provided to
Sawtooth or `ChoiceModelR` as levels. Moreover, it is important that the
value that should be interpolated needs to lie within the lower and
upper bound of `interpolate.levels`. In our case, we had 7 levels that
range from 10 to 70.

@@

In this case, the group assignment is stored in form of integers.
However, the output is the same if it is a factor or if it is labelled
data. To proof this we just quickly change `group` into a factor by
using the `factor()` function provided by the *R* Core Team (2023).

@@

Chrzan, K., & Orme, B. K. 2019. <em>Applied MaxDiff: A Practitioner's
Guide to Best-Worst Scaling</em> Provo, UT: Sawtooth Software.

@@

Gilbride, Timothy J., Lenk, Peter J., and Brazell, Jeff D. 2008. "Market
Share Constraints and the Loss Function in Choice-Based Conjoint
Analysis." <em>Marketing Science
27</em>(6):995-1011.<https://doi.org/10.1287/mksc.1080.0369>

@@

@@

Miaoulis, G., Parsons, H., & Free, V. 1990. Turf: A New Planning
Approach for Product Line Extensions. <em>Marketing Research 2</em> (1):
28-40.

@@

Sermas R (2022). *ChoiceModelR: Choice Modeling in R*. R package version
1.3.0, <https://CRAN.R-project.org/package=ChoiceModelR>.

>>>>

The goal of `validateHOT` is to provide a tool to validate the results
of your validation task (also known as holdout task; short form
**HOT**), simulating markets with the results of your MaxDiff, CBC, and
ACBC, and lastly to convert the raw logit scores of your MaxDiff, CBC,
and ACBC into scores that are easier to communicate. This package is
especially relevant for the [Sawtooth
Software](https://sawtoothsoftware.com/) community who would like to
report their analysis in *R* for open science purposes. However, it also
works perfectly with the great package `ChoiceModelR`(Sermas, 2022).
This package is extremely useful for practitioners, who would like to
run the analyses in an open source software. The ultimate goal of these
preference measurement techniques is to predict future behavior (Green &
Srinivasan, 1990). Therefore, it is essential for both academics and
practitioners to ensure that the collected data is valid and can also
predict outside tasks that were not included in the estimation of the
utility scores.

Although commercial studies often do not include a validation/holdout
task (Yang et al., 2018), it is highly recommended to do so (Orme, 2015;
Rao, 2014). Validation/holdout tasks do not only check whether your data
is valid but can also help to test different models. `validatHOT`
provides some of the relevant metrics to test the performance of the
data in predicting a validation/holdout task.

@@

👉🏾 <u>**What you need to provide**</u>: <br> After collecting your data
and running your initial Hierarchical Bayes models, you can turn to
`validateHOT` and test how good your model predicts choices in the
validation/holdout task. Therefore, you only have to read in your raw
utility scores as well as the actual choice of your validation/holdout
task. You can use the `merge()` function provided by the *R* base
package to do so (R Core Team, 2023). Afterward, you can read in your
data file and enjoy `validateHOT`. We provide a short tutorial in the
markdown, for a more detailed tutorial, please see the vignette that
comes along with `validateHOT`.

@@

<li>validation metrics</li>

@@

<li><code>hitrate()</code>: creates the Hit Rate (correctly predicted
choices) of your validation/holdout task. The output will contain the
percentage of how many choices were correctly predicted including the
standard error, the chance level in your validation task
($\frac{1}{alternatives}$) in percentage, the number of correctly
predicted participants' choices as well as the number of
observations.</li>

<li><code>kl()</code>: Kullback-Leibler-Divergence measures the
divergence between the actual choice distribution and the predicted
choice distribution (Ding et al., 2011; Drost, 2018). Output provides
both divergence between predicted from observed and observed from
predicted due to the asymmetry of the Kullback-Leibler divergence.
`validateHOT` currently provides two logarithm bases: \eqn{log} and
\eqn{log{_2}}. The default is set to \eqn{log}.</li>

@@

<li><code>medae()</code>: calculates the median absolute error, which
is, compared to the mean absolute error, less affected by outliers.</li>

@@

<li><code>rmse()</code>: provides the root-mean-square error of
deviation between predicted and stated choice share.</li>

@@

We also include metrics from machine learning, i.e., the confusion
matrix (e.g., Burger, 2018). For all of the five provided functions, a
**none** alternative has to be included in the validation/holdout task.
The logic of the implemented confusion matrix is to test, for example,
whether a buy or no-buy was correctly predicted. Information could be
used to get a sense of overestimation and underestimation, respectively,
of product demand. In the table below <code>TP</code> stands for true
positives, <code>FP</code> for false positives, <code>TN</code> for true
negatives, and <code>FN</code> for false negatives (Burger, 2018; Kuhn,
2008). To translate this to the `validateHOT` logic, imagine you have a
validation/holdout task with five alternatives plus the alternative not
to buy any of those chosen. `validateHOT` now measures whether or not a
buy (participant opts for one of the five alternatives) or a no-buy
(participant opts for the none alternative), respectively, is correctly
predicted.

Please be aware that `validateHOT` applies the following coding of the
*buy* and *no-buy* alternatives. Rows refer to the observed decisions
while columns refer to the predicted ones.

|        | Buy | No-buy |
|--------|:---:|:------:|
| Buy    | TP  |   FN   |
| No-buy | FP  |   TN   |

@@

<li><code>turf()</code>: <b>T</b>(otal) <b>U</b>(nduplicated)
<b>R</b>(each) and <b>F</b>(requency) is a "product line extension
model" (Miaoulis et al., 1990, p. 29) that helps to find the perfect
product bundle based on the reach (e.g., how many participants consider
buying at least one product of that assortment) and the frequency (how
many products on average are a purchase option). <code>turf()</code>
currently provides both the *threshold* approach (<code>approach =
'thres'</code>; all products that exceed a threshold are considered,
e.g., as purchase option; Chrzan & Orme, 2019, p. 112) and the *first
choice* approach (<code>approach = 'thres'</code>; only product with
highest utility is considered as purchase option; Chrzan & Orme, 2019,
p. 111).</li>

<li><code>freqassort()</code>: Similar to <code>turf()</code>,
<code>freqassort()</code> will give you the averaged frequency, how many
products the participants will choose from a potential assortment.
Again, you have to define a <code>none</code> alternative, because
<code>freqassort()</code> uses the <i>threshold</i> approach, meaning if
the utility of one product is above the utility of <code>none</code>, it
is marked as potential purchase option. While <code>turf()</code>
calculates the reach and frequency for <b>all</b> combinations, you
specify the combination you are interested in
<code>freqassort()</code>.</li>

@@

<li><code>marksim()</code>: Runs market simulations (either share of
preference or first choice rule), including the standard error, as well
as the lower and upper confidence interval, which is calculated
according to the following formula $mean +/- 1.96 x \frac{sd}{\sqrt(n)}$
(Orme, 2020, p. 94).</li>

@@

### Converting raw utilities

<code>validateHOT</code> also provides four functions to better
interpret the scores of both (A)CBC and MaxDiff, namely:

@@

<li><code>att_imp()</code>: Converts the raw utilities of either an ACBC
or CBC into importance scores for each attribute (see, Orme, 2020, pp.
79-81)</li>

<li><code>prob_scores()</code>: Converts the raw utilities of a MaxDiff
to choice probabilities by applying the following procedures:

@@

<li>For anchored MaxDiff a slightly different formula is applied:
$\frac{exp^{U_i}}{(exp^{U_i} + a - 1)} * 100 * \frac{1}{a}$ (Chrzan &
Orme, 2019, p. 59).</li>

@@

<li><code>zc_diffs()</code>: Rescales the raw logit utilities to make
them comparable across participants (Sawtooth Software Inc, 2023, p.
343).</li>

<li><code>zero_anchored()</code></li>: Rescales the raw logits of a
MaxDiff to zero-centered diffs (Chrzan & Orme, 2019, p. 64).

@@

`validateHOT` provides five data sets that help to better explain the
functions as well as the structure of the input, especially for the
`createHOT()` function.

@@

<li><code>CBC_lin</code>: Example data set with raw utilities of a CBC
study conducted in Sawtooth. One attribute was linear-coded while the
other attributes are part-worth coded.</li>

@@

## The story behind `validateHOT`

We are teaching a preference measurement seminar for students. Often
these students did not have any prior experience (or only sparsely) with
*R*. One of the chapters in this class is about model validation by
checking holdout task and we teach this, of course, in *R* 😍. We want
to put a strong emphasis on open science and providing tools to run the
analyses and provide the code afterwards is essential. `validateHOT`
makes this process look easy.

Of course, there are other great packages which are faster in terms of
running time (i.e., `Metrics` by Hamner & Frasco, 2018), however, these
packages need some more data wrangling in order to use the appropriate
functions with the raw utilities, which might be a burden or barrier for
the one or the other.

Moreover, as Yang et al. (2018) report, commercial studies often do not
use any validation task. Again, the missing experience in *R* could be
one explanation. Since these functions are not always implemented in
other software, this might be one reason why they do not include one
simply because they do not know how to use it correctly. Having a
package to evaluate the validation/holdout task can be very beneficial
from this perspective.

@@

First, we load the package:

@@

Now imagine you included a validation/holdout task with three
alternatives plus a no-buy alternative. We specify the `data` argument
and as well as the `id`. Since we also have a *no-buy* alternative in
our validation task, we next specify the `none` argument, otherwise we
would have left it empty. Afterwards, we define each alternative with
the argument `prod.levels`. If we look back at the data frame, we can
see that the first alternative in the holdout task (`c(4, 9, 19)`) is
composed of the following attribute levels `Att1_Lev1`, `Att2_Lev1`, and
`Att3_Lev5`.

@@

> ❗ `createHOT()` takes both column names and column indexes. However,
> please be aware, if you include linear-coded or piecewise-coded, you
> **have** to provide the column indexes for the input in `prod.levels`.
> However, you can easily find out the index by either using the
> `names()` function to see all names of the data frame, or if you
> already know the name of the variable, you can use `which()` and
> `colnames()`, both functions are provided by the base package (R Core
> Team, 2023). For example, if you want to find out the column index of
> `ID`, you could run `which(colnames(CBC) == "ID")` to determine the
> column index of `ID`.

@@

Let us take a glimpse at the output, which shows the participants' total
raw utilities for each of the three alternatives that were included in
the validation/holdout task.

@@

In the next step, we would like to see how well our model (from which we
took the raw utilities) predicts the actual choices in the
validation/holdout task. First, we will run the `hitrate()` function. We
specify the `data`, the column names of the alternatives (`opts`;
remember there are three alternatives + the *no-buy* alternative), and
finally the actual choice (`choice`).

@@

Finally, let us test, how many participants would at least buy one of
the three products, assuming that this is one potential assortment we
would like to offer to our consumers. We will use the `reach()`
function. To specify the bundles we are offering we use the `opts`
argument in our function.

@@

### Example II - CBC with linear-coded attribute(s)

In a second example, we again use a *CBC*, however, this time we show
how to use `validateHOT` if one of the variables is linear-coded. All
other examples are provided in the accompanied vignette.

@@

Next, we create the validation/holdout task to evaluate it in the next
step. We use a validation/holdout task with 3 alternatives plus the
*no-buy* alternative, just as we did in the previous example. The only
difference to the previous example is that the third attribute
(`Att3_Lin`) was linear-coded.

Again, we first define data, namely the `id` as well as the `none`
alternative. Next, we define the `prod.levels` for each alternative.
Since we have one linear coded attribute, we have to specify the column
indexes instead of the column names. For example, for the second
alternative which is composed of `Att1_Lev5`, `Att2_Lev4`, and `60`
which is the value that should be interpolated. We tell `createHOT()`
that this value needs to be interpolated by specifying the `coding`
argument accordingly. This tells us that the first two attributes
(`Att1_Lev5`, `Att2_Lev4` in the case of alternative 2) are part-worth
coded (`0`) while the third attribute is linear-coded (`1`).

To interpolate the value, we have to provide `createHOT()` the
`interpolate.levels`. These **need** to be the same levels as provided
to Sawtooth or `ChoiceModelR`. Moreover, it is important that the value
that should be interpolated needs to lie within the lower and upper
bound of `interpolate.levels`. In our case, we had 7 levels that range
from 10 to 70.

@@

In this case, the group assignment is stored in form of integers.
However, the output is the same if it is a factor or labelled data. To
proof this we just quickly change `group` into a factor by using the
`factor()` function provided by the *R* Core Team (2023).

@@

Chrzan, K., and Orme, B. K. 2019. <em>Applied MaxDiff: A Practitioner's
Guide to Best-Worst Scaling</em> Provo, UT: Sawtooth Software.

@@

Gilbride, Timothy J., Lenk, Peter J., and Brazell, Jeff D. 2008. "Market
Share Constraints and the Loss Function in Choice-Based Conjoint
Analysis." <em>Marketing Science 27</em>(6):
995-1011.<https://doi.org/10.1287/mksc.1080.0369>

@@

Kuhn, Max. 2008. "Building Predictive Models in R Using the caret
Package." <em>Journal of Statistical Software 28</em>(5), 1-26.
<https://doi.org/10.18637/jss.v028.i05>.

@@

Miaoulis, G., Parsons, H., and Free, V. 1990. Turf: A New Planning
Approach for Product Line Extensions. <em>Marketing Research 2</em>(1):
28-40.

@@

Sermas, R. 2022. *ChoiceModelR: Choice Modeling in R*. R package version
1.3.0, <https://CRAN.R-project.org/package=ChoiceModelR>.

### 2023-12-21 paper/paper.Rmd
proof reading of manuals, markdown, and vignette - 231221

---
title: 'validateHOT - Validate your Holdout Task'
tags:
  - R
  - MaxDiff
  - Conjoint Analysis
  - Market Simulations
  - Predictive Validity
authors:
  - name: Joshua Schramm
    orcid: 0000-0001-5602-4632
    corresponding: True
    affiliation: 1, 2
  - name: Marcel Lichters
    orcid: 0000-0002-3710-2292
    corresponding: FALSE
    affiliation: 1, 2    
affiliations:
 - name: Chemnitz University of Technology, Germany
   index: 1
 - name: Otto von Guericke University of Magdeburg, Germany
   index: 2
citation_author: Schramm & Lichters
date: 22 October 2023
year: 2023
bibliography: paper.bib
link-citations: true
output: rticles::joss_article
journal: JOSS
---

@@

validateHOT is a practical tool for Sawtooth Software users in industry as well as academia. It provides an open source solution for a) validating a validation/holdout task and ensuring that the model has predictive validity; b) running market simulations (e.g., Total Unduplicated Reach and Frequency, hereafter TURF). Other packages, for example, Metrics [@Metrics] provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix. However, to put the Sawtooth export into the right format, the user needs some data wrangling which could pose a barrier. Moreover, there are also packages that however mainly focus on the analysis of conjoint analysis (e.g., ChoiceModelR, @ChoiceModelR, choicetools, @choicetools, logitR, @logitr, bayesm, @bayesm], etc.). To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations (especially TURF) is missing.

@@

validateHOT's functions can be categorized into four main areas, see \autoref{tab:table1}. To bring the data into the right format, users can run the \texttt{\color{purple}createHOT()} function, which creates the total utility of each alternative by applying the additive utility model [@rao2014, p. 82]. \colcod{turf()} as well as the 3 rescaling functions, however, are not dependent on \texttt{\color{purple}createHOT()}, and can be run using the raw logit scores.

| Validation metrics | Confusion matrix | Market simulations | Rescaling scores |
|:----------------:|:----------------:|:----------------:|:----------------:|
|     hitrate()      |    accuracy()    |    freqassort()    |    att_imp()     |
|        kl()        |       f1()       |     marksim()      |  prob_scores()   |
|       mae()        |   precision()    |      reach()       |    zc_diffs()    |
|      medae()       |     recall()     |       turf()       |                  |
|       mhp()        |  specificity()   |                    |                  |
|       rmse()       |                  |                    |                  |

>>>>

---
title: 'validateHOT - Validate your Holdout Task'
tags:
  - R
  - MaxDiff
  - Conjoint Analysis
  - Market Simulations
  - Predictive Validity
authors:
  - name: Joshua Benjamin Schramm
    orcid: 0000-0001-5602-4632
    corresponding: True
    affiliation: 1, 2
  - name: Marcel Lichters
    orcid: 0000-0002-3710-2292
    corresponding: FALSE
    affiliation: 1, 2    
affiliations:
 - name: Chemnitz University of Technology, Germany
   index: 1
 - name: Otto von Guericke University of Magdeburg, Germany
   index: 2
citation_author: Schramm & Lichters
date: 22 October 2023
year: 2023
bibliography: paper.bib
link-citations: true
output: rticles::joss_article
journal: JOSS
---

@@

validateHOT is a practical tool for Sawtooth Software users in industry as well as academia. It provides an open source solution for a) validating a validation/holdout task and ensuring that the model has predictive validity; b) running market simulations (e.g., Total Unduplicated Reach and Frequency, hereafter TURF). Other packages, for example, Metrics [@Metrics] provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix. However, to put the Sawtooth export into the right format, the user needs some data wrangling which could pose a barrier. Moreover, there are also packages that however mainly focus on the analysis of conjoint analysis (e.g., ChoiceModelR, @ChoiceModelR, choicetools, @choicetools, logitR, @logitr, bayesm, @bayesm, etc.). To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations (especially TURF) is missing.

@@

validateHOT's functions can be categorized into four main areas, see \autoref{tab:table1}. To bring the data into the right format, users can run the \texttt{\color{purple}createHOT()} function, which creates the total utility of each alternative by applying the additive utility model [@rao2014, p. 82]. \colcod{turf()} as well as the four rescaling functions, however, are not dependent on \texttt{\color{purple}createHOT()}, and can be run using the raw logit scores.

| Validation metrics | Confusion matrix | Market simulations | Rescaling scores |
|:----------------:|:----------------:|:----------------:|:----------------:|
|     hitrate()      |    accuracy()    |    freqassort()    |    att_imp()     |
|        kl()        |       f1()       |     marksim()      |  prob_scores()   |
|       mae()        |   precision()    |      reach()       |    zc_diffs()    |
|      medae()       |     recall()     |       turf()       | zero_anchored()  |
|       mhp()        |  specificity()   |                    |                  |
|       rmse()       |                  |                    |                  |
